---
title: "Probability homework"
output: html_notebook
---

```{r}
library(prob)
library(gtools)
```
#### 1. If I roll a six-sided dice once what is the probability of getting a number 
####    less than 5?
Sample space
S = {1,2,3,4,5,6}
Event "roll is less than 5"
Se = {1,2,3,4} Event space
**Probability = 4/6**

#### 2. How many outcomes are there in the sample space of an experiment in which we
####    toss three dice? **216**

```{r}
S <- rolldie(times = 3, makespace = TRUE)
size_S <- nrow(S)
size_S

```
#### 3.i What is the probability of getting a total of the three dice equal
####     to 5, 6 or 7? **0.1435185**
#### ii. How many outcomes lie within this event total_equals_5_6_7?  **31**

```{r}
S <- rolldie(3, makespace = TRUE)
# makespace adds the probs column
S <- addrv(space = S, FUN = sum, invars = c("X1", "X2", "X3"), name = "Total")
S
size_S <- nrow(S)
size_S

total_equals_5_6_7 <- subset(S, Total == 5 | Total == 6 | Total == 7)
total_equals_5_6_7
size_S_total_5_6_7 <- nrow(total_equals_5_6_7)
size_S_total_5_6_7

prob_total_5_6_7 <- size_S_total_5_6_7 / size_S
prob_total_5_6_7
```

#### 4. Five chairs are in a room along with five people of differing height.
####    The people sit down randomly on the chairs.
####    What is the probability that they sit in order of decreasing height,
####    left to right? **1/120**
```{r}

people <- c("height5", "height4", "height3", "height2", "height1")

chairs <- permutations(n = 5, r = 5, v = people, repeats.allowed = FALSE)

chairs
# 120 different combinations with one of those being the correct order.
# Probability 1/120
```
#### 5. I roll six dice.
####    What is the probability I see the numbers 6,5,4,3,2,1 in order on
####    the dice, left to right?
```{r}

```



#### 6. If I draw a single card from a 52 card deck, and it is black, what is
####    the probability that its suit is spades? **0.5**
```{r}
# Sample space reduced to 26 as we know the card is black.
# 13 out of the 26 cards are spades so the probability = 13/26 = 0.5
```

