---
title: "Model Building Homework"
output: html_notebook
---


# MVP

You are given a set of data on housing sale prices for the last few years in King County (near Seattle) between May 2014 and May 2015. Have a look at the variable definitions on the [Kaggle page](https://www.kaggle.com/harlfoxem/housesalesprediction)

```{r}
library(readr)
kc_house_data <- read_csv("kc_house_data.csv")
```


* Tidy up the data ready for regression:

    * You might like to think about removing some or all of `date`, `id`, `sqft_living15`, `sqft_lot15` and `zipcode` (`lat` and `long` provide a better measure of location in any event).

```{r}
str(kc_house_data)
```
    
```{r}
kc_house_data$id <- NULL
```
```{r}
kc_house_data$sqft_living15 <- NULL
```
```{r}
kc_house_data$sqft_lot15 <- NULL
```
```{r}
kc_house_data$zipcode <- NULL
```

  
    * Have a think about how to treat `waterfront`. Should we convert its type?

```{r}
unique(kc_house_data$waterfront)
```
```{r}
kc_house_data$waterfront <- as.logical(kc_house_data$waterfront)
```
```{r}
unique(kc_house_data$waterfront)
```


    * We converted `yr_renovated` into a `renovated` logical variable, indicating whether the property had ever been renovated. You may wish to do the same.
```{r}
library(tidyverse)
unique(kc_house_data$yr_renovated)
```
```{r}
# addin a true/false renovated column 
kc_house_data$renovated <- ifelse(kc_house_data$yr_renovated > 0, TRUE, FALSE)
```

```{r}

```

    * Have a think about how to treat `condition` and `grade`? Are they interval or categorical ordinal data types?
```{r}
# condition and grade are categorical ordinal data types.
unique(kc_house_data$condition)
unique(kc_house_data$grade)
```
```{r}
library(fastDummies)
kc_house_data <- dummy_cols(kc_house_data, select_columns = c("condition", "grade"), remove_first_dummy = TRUE)
glimpse(kc_house_data)
```


* Check for aliased variables using the `alias()` function (this takes in a formula object and a data set). [**Hint** - formula `price ~ .` says 'price varying with all predictors', this is a suitable input to `alias()`]. Remove variables that lead to an alias.

```{r}
kc_house_data$date <- NULL
```

```{r}
alias(price ~ ., kc_house_data)
```
```{r}
?alias
```
Looked up hep and online but not really sure what to so with the result of this.



* Systematically build a regression model containing up to **four** main effects (remember, a main effect is just a single predictor with coefficient), testing the regression diagnostics as you go

```{r}
library(tidyverse)
library(modelr)
library(ggiraphExtra)
library(GGally)
```


```{r}
library(psych)
```

```{r}
pairs.panels(kc_house_data[c("price", "bedrooms", "bathrooms", "sqft_living", "sqft_lot", "floors", "view", "condition", "sqft_above", "sqft_basement")])
```

```{r}
mod1a <- lm(price ~ sqft_living, data = kc_house_data)
summary(mod1a)

mod1b <- lm(price ~ sqft_above, data = kc_house_data)
summary(mod1b)
```
1a better (sqft_living) r sqrd 0.4929

```{r}
par(mfrow = c(2,2))
plot(mod1a)
plot(mod1b)
```

```{r}
kc_house_data_remaining_resid <- kc_house_data %>%
  add_residuals(mod1a) %>%
  select(-c("price", "sqft_living"))

# manually removed "price" and "sqft living"
pairs.panels(kc_house_data_remaining_resid[c("bedrooms", "bathrooms", "sqft_lot", "floors", "view", "condition", "sqft_above", "sqft_basement")])
```
Try bathrooms and the sqft_above
```{r}
mod2a <- lm(price ~ sqft_living + bathrooms, data = kc_house_data)
summary(mod2a)
mod2b <- lm(price ~ sqft_living + sqft_above, data = kc_house_data)
summary(mod2b)
```
not much in it mod2b best r sqrd 0.4933
```{r}
par(mfrow = c(2,2))
plot(mod2a)
plot(mod2b)
```

mod2b better (marginally)
in both items are appearing in Cooks distance so some outliers having a
disproportionate effect.

```{r}
anova(mod2a, mod2b)
```

```{r}
kc_house_data_remaining_resid <- kc_house_data %>%
  add_residuals(mod2b) %>%
  select(-c("price", "sqft_living", "sqft_above"))

#manually removed "price", "sqft_living", "sqft_above"
pairs.panels(kc_house_data_remaining_resid[c("bedrooms", "bathrooms", "sqft_lot", "floors", "view", "condition", "sqft_basement")])


```
Bathrooms and sqft_basement next to try

```{r}
mod3a <- lm(price ~ sqft_living + sqft_above + bathrooms, data = kc_house_data)
summary(mod3a)
mod3b <- lm(price ~ sqft_living + sqft_above + sqft_basement, data = kc_house_data)
summary(mod3b)
```
no difference in R squrd figure of 0.4933 for both.

```{r}
anova(mod3a, mod3b)
```








    * splitting datasets into numeric and non-numeric columns might help `ggpairs()` run in manageable time, although you will need to add either a `price` or `resid` column to the non-numeric dataframe in order to see its correlations with the non-numeric predictors.

<details>
<summary>**Hints**</summary>
```{r, eval=FALSE}
houses_tidy_numeric <- houses_tidy %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- houses_tidy %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- houses_tidy$price

ggpairs(houses_tidy_numeric)
ggpairs(houses_tidy_nonnumeric)
```
and the same in subsequent rounds of predictor selection with the `resid` column.<br><br>
Remember, if you are not sure whether including a categorical predictor is statistically justified, run an `anova()` test passing in the models with- and without the categorical predictor and check the p-value of the test.
</details>







# Extensions

* Consider possible interactions between your four main effect predictors and test their effect upon $r^2$. Choose your best candidate interaction and visualise its effect. 

* Calculate the relative importance of predictors from your best $4$-predictor model (i.e. the model without an interaction). Which predictor affects `price` most strongly?



```{r}

```

