---
title: "Anscombe Quartet Homework"
output: html_notebook
---

summary(anscombe)
```{r}
str(anscombe)
```

**x1**
```{r}
mean(anscombe$x1)
var(anscombe$x1)
```
**x2**
```{r}
mean(anscombe$x2)
var(anscombe$x2)
```
**x3**
```{r}
mean(anscombe$x3)
var(anscombe$x3)
```
**x4**
```{r}
mean(anscombe$x4)
var(anscombe$x4)
```
**y1**
```{r}
mean(anscombe$y1)
var(anscombe$y1)
```
**y2**
```{r}
mean(anscombe$y2)
var(anscombe$y2)
```
**y3**
```{r}
mean(anscombe$y3)
var(anscombe$y3)
```
**y4**
```{r}
mean(anscombe$y4)
var(anscombe$y4)
```

Looking just at the `mean` and `variance` values for each variable it appears 
that all the "x" values `x1, x2, x3` and `x4` appear very similar.

Likewise all the "y" values `y1, y2, y3` and `y4` appear very similar, again, 
only judging by the `mean` and `variance` values.

I'm sure this won't turn out to be the case when we visualise the data.

```{r}
anscombe
```
Using an enhanced pairs.panel function from the "psych' library to give a
very quick indication across all of the `x` values and then all of the `y`
values.

```{r}
library(psych)
pairs.panels(anscombe[c("x1", "x2", "x3", "x4")])
pairs.panels(anscombe[c("y1", "y2", "y3", "y4")])
```

The above plot and the boxplot below shows that some datasets do vary.
x1 - x3 look to have similar properties in both plots. However, x4, y1 and y2
are clearly different. y3 and y4 look similar but do have differences.
```{r}
boxplot(anscombe)
```

```{r}
IQR(anscombe$x1)
e1071::skewness(anscombe$x1)
```

```{r}
IQR(anscombe$x2)
e1071::skewness(anscombe$x2)
```

```{r}
IQR(anscombe$x3)
e1071::skewness(anscombe$x3)
```

```{r}
IQR(anscombe$x4)
e1071::skewness(anscombe$x4)
```
```{r}
anscombe$x4
```

```{r}
IQR(anscombe$y1)
e1071::skewness(anscombe$y1)
```
```{r}
IQR(anscombe$y2)
e1071::skewness(anscombe$y2)
```
```{r}
IQR(anscombe$y3)
e1071::skewness(anscombe$y3)
```
```{r}
IQR(anscombe$y4)
e1071::skewness(anscombe$y4)
```


Not sure I trust the correlation figures of 0.82 for all of the `(x,y)` 
relationships.
```{r}
psych::pairs.panels(anscombe[c("x1", "y1", "x2", "y2", "x3", "y3", "x4", "y4")])
```
Each of the plots of the `(x,y)` relationship between the two sets of data shows
a different pattern on the scatterplot. `(x1, y1)` looks like you could fit a line
through some of the points although some points are further away.
```{r}
plot(x1 ~ y1, data = anscombe)
```
This plot shows a curved relation ship between the two variables.
```{r}
plot(x2 ~ y2, data = anscombe)
```

This appears to show a straight-line relationship between the two variables
but there is one significant outlier that does not fit with this.
```{r}
plot(x3 ~ y3, data = anscombe)
```

Again, appears to show a stright-line relationship with one significant outlier.
```{r}
plot(x4 ~ y4, data = anscombe)
```
**Correlation coefficients for each `(x,y)` set**
```{r}
cor(anscombe$x1, anscombe$y1)
cor(anscombe$x2, anscombe$y2)
cor(anscombe$x3, anscombe$y3)
cor(anscombe$x4, anscombe$y4)
```

